---
title: "DATA 643 Project 1"
author: "Burton Machado"
date: "June 19, 2016"
output: pdf_document
---

**We will build a recommender system to recommend books to users using user-user collaborative filtering.**

**Following is the data that is used. This is included in the attached Project_1.csv file.**

Reviewer,Learn Java,Oracle PL SQL,R in Action,Learning Python,Learn MongoDB,Scala in Action

John,4,,5,,3,

James,5,,4,3,4,3

Debbie,3,2,,,3,2

Sarah,4,3,3,,3,

Steven,3,,,3,,4

Sergey,4,,4,,,3

Brad,,3,,4,,4

Allison,5,,5,,3,4

Megh,,4,,4,,3

Alan,4,,4,,3,

**As shown above, we have 10 users and 6 books to recommend. This is a artificially created "toy" data set.**

We will use the dataset and create the intermediate data structures to build our user-user recommender system. Specifically we will build the following intermediate data structures and print them.

1) The initial data frame by reading the csv data file.

2) The data frame with added user means and book means.

3) A data frame containing  the mean centered values.

4) The mean centered data frame with the distance and user means.

5) The similarity matrix.

We can then use these data structure to predict the missing ratings of users.

As an example, we will calculate the predicted rating for Sarah for the book 'Scala in Action'

Please note that a datafarme will be broken into two for display purposes peices if it is too wide to display on the page.


```{r warning=FALSE, echo=FALSE}
library(knitr)
library("recommenderlab")
```

```{r warning=FALSE}

# Please set the working directory to the location of  the 'Project_1.csv' file.
setwd('/Users/burton/001-Semester_05_CUNY/643_Recommender_Systems/Week_01')
# Create a data frame from the csv file.
df <- read.csv('Project_1.csv', header = TRUE, stringsAsFactors = FALSE)
# We use kable function from knitr to format and print a dataframe. If a dataframe is too wide to 
# fit on the page, we split it into multiple tables.
kable(df[,1:4])
kable(df[,5:7])
# Here we will calculate the user means and bind them to the data frame
userMeans <- c()
for(i in 1:nrow(df)){
  rowWithoutNA <- df[i, 2:ncol(df)][!is.na(df[i, 2:ncol(df)])]
  userMeans <- append(userMeans, round(mean(rowWithoutNA), 2))
}
df <- cbind(df, userMeans)

# Here we will calculate the Book means and bind them to the data frame
bookMeans <- c(NA)
for(i in 2:ncol(df)){
  colWithoutNA <- df[,i][!is.na(df[,i])]
  bookMeans <- append(bookMeans, round(mean(colWithoutNA), 2))
}
df <- rbind(df, bookMeans)
df[nrow(df),1] <- 'bookmeans'
# Print the dataframe along with the user-means and book-means
kable(df[,1:4])
kable(df[,5:8])

# We now calculate the mean-centered values and create another dataframe.
df_Mean_Centered <- df[,2:ncol(df)]-df$userMeans
df_Mean_Centered <- df_Mean_Centered[,1:ncol(df_Mean_Centered)-1]
df_Mean_Centered <- df_Mean_Centered[1:nrow(df_Mean_Centered)-1,]
# Print the mean-centered dataframe in a table.
kable(df_Mean_Centered)

# Calculate the distances
distance <- c()
for( i in 1:nrow(df_Mean_Centered)){
  distance <- append(distance, sum(df_Mean_Centered[i,]**2, na.rm = TRUE)**.5)
}
#Create mean centered data frame with the distance
df_Mean_Centered_and_Distance <- cbind(df_Mean_Centered, distance)
df_Mean_Centered_and_Distance <- cbind(df_Mean_Centered_and_Distance, userMeans=df$userMeans[1:length(df$userMeans)-1])

#Construct the similarity matrix
similarity_Matrix <- data.frame(names=df[1:nrow(df)-1,1])
for(i in 1:nrow(df_Mean_Centered_and_Distance)){
  temp <- c()
  for(j in 1:nrow(df_Mean_Centered_and_Distance)){
    product <- sum(df_Mean_Centered_and_Distance[i,1:6]*df_Mean_Centered_and_Distance[j,1:6],na.rm = TRUE)/
      (df_Mean_Centered_and_Distance$distance[i]*df_Mean_Centered_and_Distance$distance[j])
    temp <- append(temp, product)
  }
  similarity_Matrix <- cbind(similarity_Matrix, temp)
}
names(similarity_Matrix)<-c('names', as.character(similarity_Matrix[,1]))
  
df_Mean_Centered_and_Distance <- cbind(Reviewer=df[1:10,1],df_Mean_Centered_and_Distance)
#df_Mean_Centered_and_Distance
kable(df_Mean_Centered_and_Distance[,1:5])
kable(df_Mean_Centered_and_Distance[,6:9])
#similarity_Matrix
kable(similarity_Matrix[,1:6])
kable(similarity_Matrix[,7:11])
  
  # We will Rate scala in action for Sarah.
  
# The closest two users to Sarah are James(0.5520524) and Allison(0.4787136)
  
similarity_Sarah <- sort(similarity_Matrix$Sarah,decreasing = TRUE)
  
  
# We take the sum of ratings of James and Allison multiplied by their similarity to Sarah, 
# then divide by the sum of their similarities to Sarah.
  
rating_Sarah_Scala <- similarity_Sarah[2]*df_Mean_Centered$Scala.in.Action[2] + similarity_Sarah[3]*df_Mean_Centered$Scala.in.Action[8] + df_Mean_Centered_and_Distance$userMeans[4]
```

#**The predicted rating for Sarah for the book 'Scala in Action' is `r round(rating_Sarah_Scala,0)`**

**Now we will try to predict rating for Sarah for the book 'Scala in Action' using the recommenderlab package.**

```{r}
mm<-as.matrix(df[1:nrow(df),2:ncol(df)])
rr<-as(mm, "realRatingMatrix")

rec=Recommender(rr[1:10],method="UBCF", param=list(normalize = "Z-score",method="Cosine",nn=5))
recom <- predict(rec, rr[1:10], type="ratings")
recom_list <- as(recom, "list")
```

#**The predicted rating for Sarah for the book 'Scala in Action' using the recommenderlab package is  `r round(recom_list[[4]][[2]],0)`. The ratings from our code and from the package match.**

Note that in the list above Sarah is the 4th user and 'Scala in Action' is the 2nd predicted entry, hence the indexes 4,2

The contents of the list recom_list are displayed below.

```{r}
as(recom, "list")

```
